Climate indicators

### 2023-02-22 DRAFT_climate_indicators.Rmd
mean per year

@@

```{r}
tif = system.file("tif/L7_ETMs.tif", package = "stars")
t1 = as.Date("2018-07-31")
x = read_stars(c(tif, tif, tif, tif), along = 
                  list(time = c(t1, t1+1, t1+2, t1+32)), 
               RasterIO = list(nXOff = c(1), 
                               nYOff = c(1), 
                               nXSize = 50, 
                               nYSize = 50, 
                               bands = c(1:6)))
```

@@

, and four dimensions

@@

X and y area the coordinates. Band is

@@

an integer.

And time is

@@

Four dates covering three months in 2018.

@@

The workflow is as such:

@@

### Filter by dates

>>>>

### Additional resources

[Stars package](https://r-spatial.github.io/stars/)

[R as a GIS for economists](https://tmieno2.github.io/R-as-GIS-for-Economists/stars-basics.html) chapter 7

@@

```{r}
tif = system.file("tif/L7_ETMs.tif", package = "stars")
t1 = as.Date("2018-05-31")
x = read_stars(c(tif, tif, tif, tif), along = 
                  list(time = c(t1, t1+1, t1+50, t1+100)), 
               RasterIO = list(nXOff = c(1), 
                               nYOff = c(1), 
                               nXSize = 50, 
                               nYSize = 50, 
                               bands = c(1:6)))
```

@@

I can rename it like this
```{r}
x <- setNames(x, "Attribute_A")
```

And I can add another dummy attribute.

```{r}
x <- x %>%
  mutate("Attribute_B" = Attribute_A/2)
```

The dummy data also has four dimensions

@@

X and y area the coordinates. Band is an integer:

@@

And time is four dates covering four months in 2018:

@@

#### Ecosystem map

Coming soon ....

@@

The general workflow is like this:

@@

I will try to use [parallelization](https://tmieno2.github.io/R-as-GIS-for-Economists/EE.html) to speed things up. Also, masking to ecosystem time earlier in the workflow might prove smart.

### Calculate reference values
#### Import

I need to code a for loop here for importing the .nc files.

But for then selecting the attribute, simply do like this:

```{r}
names(x)
```
```{r}
x_sub <- x %>% 
  select(Attribute_A)
names(x_sub)
```

#### Filter data by dates

Say we want to calculate the mean summer temperature. We then want to exclude the data for the times that is not withing our definition of summer.

Let's try it on the dummy data. Remember this data had three time steps.
```{r}
st_get_dimension_values(x, "time")
```

Our data has 365 time steps for each year.

First I define my start and end dates. I want to keep only the summer data, defined as jun - aug.

```{r}
start <-  ymd("2018-06-01")
end <-  ymd("2018-08-31")
summer <- interval(start, end)
```

Then I filter

```{r}
x_aug <- x_sub %>%
  filter(time %within% summer)
st_get_dimension_values(x_aug, "time")
```

#### Aggregate across time within a year

For this example I want to calculate the mean temperature over the summer.
I therefore need to aggregate over time. Here is how:

```{r}
x_summerMean <- x_aug %>%
  filter(band ==1) %>% # for our real data we don't need this filter
  st_apply(1:2, mean)

dim(x_summerMean)
```

The time (and band) dimension is now gone as we have aggregated accoss it.

We can also use the aggregate function to produce a data cube with lower temporal resolution, for example per year:

```{r}
x_summerMean_v2 <- x_aug %>%
  aggregate(by = "year", mean)
```

Actually, this second method is better because we don't ned to filter out additional dimensions, and it retains the attribute name (the first method replaced it with the function name _mean_).
